Hemodynamic responses of the Parahippocampal Place Area to spatial cues in a movie and its audio-description

### 2020-11-11 paper/main.tex
correction of latex syntax
restructure quotes of contextual associations
abstract, adjustments to intro & results

\begin{abstract}
% < 170 words new analysis of existing data of interest to a broad section of
% our audience highlighting innovative examples of data reuse may be used to
% present compelling new findings & conclusions derived from published.
Intro: PPA as classic example for an visual area; Methods: fMRI data
studyforrest data set, general linear model; Results: significant clusters in
PPA, RSC, individual results in x of 15 subjects; Conclusions: results suggest
generalizability to spoken language and is restricted to processing visual
stimuli exclusively. \end{abstract}

@@

% summary if literature review
In summary, studies suggest that the PPA does preferably but not exclusively
respond to visually presented spatial information.
% three things in common
Nevertheless, the reviewed studies usually had three things in common.
% designed stimulus set
First, they employed a small sets of carefully chosen stimuli. Further, stimuli
were conceptualized to control confounding variables (e.g. color, luminance,
size, sentence length), a process that is often not reported in
detail.\todo{really? check that}
% block-design & task
Second, studies used block-designed paradigms to maximize detection power.
Further, they usually employed explicit perceptual judgment to keep participants
paying attention to the stimuli.
% averaging
Third, the reviewed studies averaged results across 7-11 subjects to improve
signal-to-noise ratio (SNR).
% disadvantages
Conceptualized stimuli, block-design experiments and the averaging approach have
their merits but also disadvantages.
% designed stimulus set
First, small sets of conceptualized stimuli presented on experimental blocks
poorly resembles our rich and multidimensional environment that our brains are
accommodated to [sonkusare2019naturalistic].
% block-design & task
Second, study participants can and will become aware about which perceptual
process is examined in the study by merely being exposed to the experimental
paradigm.
% inference: validity is poor
Thus, small sets of conceptualized stimuli and block-design experiment lack
external validity \citep{westfall2016fixing} as well as ecological validity
\citep{hasson2004intersubject}.
% disadvantage of averaging
Lastly, averaging data across subjects increases the SNR but does not
characterize brain functions at an individual level, a prerequisite for the
application of brain imaging methods in individual diagnostics (cf.
\cite{dubois2016building, eickhoff2020towards}).

% Open research question
In this study, we investigate if increased activation in the PPA correlates with
incidental ``spatial information'' in the audio-description of the Hollywood
movie ``Forrest Gump'' (i.e. the audio-only variant of the movie) and, for
cross-modal comparison, in the (actual) audio-visual movie (itself).
% we are different
Since we did not design the given naturalistic stimuli (see
\citep{hamilton2018revolution, hasson2008neurocinematics,
sonkusare2019naturalistic} for reviews) for optimal localization performance of
the PPA, we operationalize the perception of landscapes and landmarks (compared
to e.g. faces and objects) differently from the reviewed studies.
% operationalization via annotation
Here, we annotated words spoken by the narrator of the audio-description as well
as cuts and depicted locations in the movie.
% public data
We then analyze publicly available data
(\href{http://www.studyforrest.org}{studyforrest.org}) from the same subjects
that listened to the audio-description and watched the movie during fMRI
scanning.
% first: group level AO
First, general linear model (GLM) contrasts based on the audio-description's
annotation are used to test if we could isolate the PPA with an auditory
stimulus on a group level.
% second: group level AV
Second as conceptual/cross-modal control, we compare results of the
audio-description to results based on the annotation of the movie.
% individual results
Lastly focussing on the audio-description, we compare results on the level of
individual subjects to results of a block-designed visual localizer experiment
done with the same individuals \citep{sengupta2016extension}.
% results
\todo[inline]{following paragraph will be similar to abstract, intro to
discussion and conclusion}
Our results demonstrate that (a), (b), (c).

@@

\todo[inline]{I consider tables of significant clusters for all contrasts as
overkill. Maybe in ``supporting material'' or reference to dataset?  all sign.
clusters of all contrasts are in the comments}

@@

@@

@@

% all eight AO PPA contrasts
In order to test the robustness of our findings, we created seven additional
$t$-contrasts for the AO stimulus (see Table \ref{tab:ao-contrasts}).
% results
Six of the overall eight AO contrasts show significant bilateral clusters in
anterior regions of the group PPA overlap (see Figure
\ref{fig:stability-slices}.\todo{Discussion: discuss that contrasts make ``less
and less'' sense}
% statement/answer
Results suggest that our findings do not depend on one specific contrasts but
are robust across differently designed contrasts.
% question
\todo[inline]{concluding statement often feels like interpretation/discussion}

% precuneus = retrosplenial cortex ?
\todo[inline]{significant clusters in other regions than PPA could or should (?)
also be reported? Following is a short list:}
% ventral precuneus = retrosplenial cortex?
7 contrasts: bilateral clusters in RSC (and ventral precuneus);
% transverse occipital sulcus = occipital place area
6 contrasts (4 bilat.): transverse occipital sulcus (= OPA?);\todo{AO: clusters
are more lateral; AV: clusters are more medial}
% confounds
4 contrasts: right superior temporal gyrus; 3 contrasts: left superior temporal
gyrus; 3 contrasts: some (smallish) prefrontal clusters

% AO negative control contrasts
\todo[inline]{control contrast at moment of cuts but in the audio-description
(contrasts 9-13) could be, if at all, mentioned here already; at the moment
negative control contrasts come last combined for AO \& AV at the end}

@@

% AV intro
Second as conceptual/cross-modal control of the group average results, we
analyzed data from the audio-visual movie by contrasting different kinds of
movie cuts independent(ly) from the frames' visual content following the cuts.
% AV primary contrast
The movie's primary $t$-contrast that compared cuts to an unfamiliar setting
with cuts within familiar setting (\texttt{vse\_new > vpe\_old}) yielded three
significant clusters (see Table \ref{tab:res-av-group1}, Figure
\ref{fig:group-slices}).
% PPA (others the clusters extend into).
One cluster spans across the midline and comprises (reported from posterior to
anterior), parts of the intracalcarine and cuneal cortex, the lingual gyrus and
retrosplenial cortex, the occipital and temporal fusiform gyrus and the
parahippocampal cortex in both hemispheres.
% LOC
Two additional homologous clusters are located in the superior lateral occipital
cortex (left bigger than right).

@@

% all eight AV PPA contrasts
Here again in order to test the robustness of our findings, we created four
additional $t$-contrasts for the AV stimulus (see Table \ref{tab:av-contrasts}).
%% results: PPA
All of the overall five AV contrasts yielded significant bilateral clusters (in
that overlap with the PPA group mask/overlap [biggest overlap in posterior part
of group ROI = temporal occipital fusiform cortex] (see Figure
\ref{fig:stability-slices}).
% lingual gyrus
In three contrasts, these homologous clusters extend into the lingual gyri.
% RSC
Further, three contrasts show significant homologous clusters in the RSC.
% LOC
Lastly, four contrasts show significant homologous clusters in lateral occipital
cortices.
% concluding statement
Here again, we can show that results can in principle be accomplished by varying
the design of the contrasts.
% question
\todo[inline]{concluding statement = interpretation/discussion?}

\subsection{negative controls}

\todo[inline]{these contrasts are listed in the methods section. Shall results
of these contrasts be reported here (at all) as well?}

% AO stimulus
AO stimulus (using annotation of movie cuts) all null results except: contrast 9
(vse\_new > pe\_old): right RSC
% correlation of these nouns with cuts
(cf. Figure \ref{fig:reg-corr}: nouns that cue change to a new setting are
``somewhat'' correlated with cuts to a new setting; nouns that cue change to a
familiar setting are somewhat correlated with cuts to a familiar setting).

% AV stimulus
AV stimulus. All contrasts using ``no cut condition'' > x yield null results.
The contrast that uses the narrator's nouns but in the AV stimulus (se\_new >
se\_old): right PPA, bilateral sup. lat. occ. cortex; LOC (l/r), right superior
parietal lobe; here again: these kind of nouns and these kind of cuts are
correlated (cf. Figure \ref{fig:reg-corr}).

@@

\citep{baldassano2013differential} regarding DMN: the set of regions connected
more strongly to anterior PPA is strikingly similar to the Default Mode Network
(DMN) (Buckner et al., 2008; Fox et al., 2005; Raichle et al., 2001), which is
known to include the parahippocampal region.
%
The only portions of the DMN that do not show differential connectivity to
anterior PPA in our data are the rostral portion of the posterior cingulate
cortex (PCC) and the superior frontal cortex.  DMN has been implicated in a
large number of internally-focused tasks, one of its key roles involves
autobiographical memory (Buckner et al., 2008).  Models of recognition memory
have previously identified parahippocampal cortex as primarily encoding spatial
context information (Eichenbaum et al., 2007).
%
Data from Aminoff et al. (2007) has suggested that an anterior portion of PPA
may be involved in recall based on spatial context.
%
Our results are consistent with anterior PPA playing a more central role in
memory than posterior PPA, given anterior PPA's connectivity with the DMN. The
fact that anterior PPA had a lower sensitivity to our abstract object stimuli
does not necessarily im ply that this region does not use object information.
%
Previous work has shown that PPA responds to objects that have spatial
associations (Aminoff et al., 2007), are space-defining (Mullally and Maguire,
2011), and are navigationally-relevant (Janzen and Van Turennout, 2004).
%
These types of responses require spatial memory and cannot be based purely on
visual features like object shape.#
%
If anterior PPA is involved in processing spatial context, then space-defining
or navigationally-relevant objects could activate anterior PPA more strongly
than our abstract objects, which were unfamiliar and provided no sense of
context or orientation.
%
Further experiments will be required to determine what type of object-related
information is used in this region.

@@

% contextual associations in anterior vs. posterior
\citep{aminoff2006parahippocampal}: We propose that the general function of the
PHC can be thought of as processing of contextual associations. A subregion of
the posterior PHC is particularly optimized to process spatially organized
contextual associations, whereas more anterior regions of the PHC are optimized
to process contextual associations in other, non-spatial domains.
% posterior PPA for visual visual associative information
Spatial contextual associations are a dominant property of places and scenes,
and thus scenes consistently activate this posterior subregion of the PHC (e.g.,
PPA; Epstein, 1998). We do not argue that the PHC is not sensitive to spaces,
places, and spatial information, but rather that it is sensitive to spatial
stimuli because of the associative information (space-related associations) that
such stimuli evoke.

% Aminoff regarding contextual processing in the PHC and beyond
\citep{aminoff2006parahippocampal} regarding contextual associations processing
in PHC and beyond: We compared contextual processing of objects that are
strongly related to a specific place (e.g., an oven is found in a kitchen) with
contextual processing of objects that are highly associated with a context but
not a specific place (e.g., a birthday cake is found at a birthday party, but
the party can take place anywhere, at home, in a restaurant, at a park).
Activity elicited by these two types of contextual objects was compared with the
activity elicited for objects that were only weakly associated with many
contexts (e.g., a lamp). Differential activity was found in the PHC and the RSC,
replicating the previous findings (Bar \& Aminoff, 2003, The
parahippocampal\dots).
% results
\citep{aminoff2006parahippocampal} Activity within the PHC, however, was
distributed along an anterior–posterior axis, with activity related to spatial
contexts focused in the posterior part of the PHC and activity evoked by
non-spatial contexts focused in the more anterior part of the PHC (Bar \&
Aminoff, 2003, Cortical analysis of visual context; Aminoff et al., 2007, The
parahippocampal cortex mediates spatial and nonspatial associations).

>>>>

\begin{abstract}
% < 170 words new analysis of existing data of interest to a broad section of
% our audience highlighting innovative examples of data reuse may be used to
% present compelling new findings & conclusions derived from published.  intro
Neuroimaging studies have parcellated the human brain into distinct functional
regions.
% PPA
The ``parahippocampal place area'' is classically considered to be a higher
visual area as increased hemodynamic activity in the PPA correlates/is
correlated with the perception of static pictures of landscapes (or landmarks)
compared to pictures of faces or objects.
% What we do
In the current study, we exploratorily tested if spatial information embedded in
the audio-description of a Hollywood movie (i.e. the movie's soundtrack with an
additional narrator) the audio-visual movie can be used to isolate the PPA on a
group average level and in individual subjects.
%
We use publicly available functional magnetic resonance imaging data
(studyforrest.org) to build canonical general linear model (GLM) $t$-contrasts
and compare current results to results that we gained from a dedicated visual
localizer experiment conducted with the same 14 participants.
% compare to visual localizer experiment group results
On a group average level, our results show significantly increased activation in
the PPA but restricted to its anterior part [as well as in the retrosplenial and
lateral occipital cortex] of both hemispheres.
% individual results
On an individual level, results show significant bilateral clusters in the
anterior PPA of nine subjects and significant unilateral clusters in two
subjects.
% conclusion: generalizability
Results suggest that increased activation in the PPA during the perception of
static pictures generalizes to the perception of auditorily presented spatial
information.
% conclusion: PPA has subregions
Lastly, our results provides further evidence that the PPA can be divided into
functional subregions.\todo{kinda bold statements; this depends on the data \&
discussion}\end{abstract}

@@

% summary if literature review
In summary, studies suggest that the PPA does preferably but not exclusively
respond to visually presented spatial information.
% three things in common
Nevertheless, the reviewed studies usually had three things in common.
% designed stimulus set
First, they employed a small sets of carefully chosen stimuli. Further, stimuli
were conceptualized to control confounding variables (e.g. color, luminance,
size, spatial frequencies, sentence length), a process that is often not
reported in detail.\todo{they usually just say ``we made the black \& white}
% block-design & task
Second, studies used block-designed paradigms to maximize detection power.
Further, they usually employed explicit perceptual judgment to keep participants
paying attention to the stimuli.
% averaging
Third, the reviewed studies averaged results across 7-11 subjects to improve
signal-to-noise ratio (SNR).
% disadvantages
Conceptualized stimuli, block-design experiments and the averaging approach have
their merits but also disadvantages.
% designed stimulus set
First, small sets of conceptualized stimuli presented on experimental blocks
poorly resembles our rich and multidimensional environment that our brains are
accommodated to [sonkusare2019naturalistic].
% block-design & task
Second, study participants can and will become aware about which perceptual
process is examined in the study by merely being exposed to the experimental
paradigm.
% inference: validity is poor
Thus, small sets of conceptualized stimuli and block-design experiment lack
external validity \citep{westfall2016fixing} as well as ecological validity
\citep{hasson2004intersubject}.
% disadvantage of averaging
Lastly, averaging data across subjects increases the SNR but does not
characterize brain functions at an individual level, a prerequisite for the
application of brain imaging methods in individual diagnostics (cf.
\cite{dubois2016building, eickhoff2020towards}).

% Open research question
In this study, we investigate if increased activation in the PPA correlates with
incidental ``spatial information'' in the audio-description of the Hollywood
movie ``Forrest Gump'' (i.e. the audio-only variant of the movie) and, for
cross-modal comparison, in the (actual) audio-visual movie (itself).
% we are different
Since we did not design the given naturalistic stimuli (see
\citep{hamilton2018revolution, hasson2008neurocinematics,
sonkusare2019naturalistic} for reviews) for optimal localization performance of
the PPA, we operationalize the perception of landscapes and landmarks (compared
to e.g. faces and objects) differently from the reviewed studies.
% operationalization via annotation
Here, we annotated words spoken by the narrator of the audio-description as well
as cuts and depicted locations in the movie.
% public data
We then analyze publicly available data
(\href{http://www.studyforrest.org}{studyforrest.org}) from the same subjects
that listened to the audio-description and watched the movie during fMRI
scanning.
% first: group level AO
First, general linear model (GLM) contrasts based on the audio-description's
annotation are used to test if we could isolate the PPA with an auditory
stimulus on a group level.
% second: group level AV
Second as conceptual/cross-modal control, we compare results of the
audio-description to results based on the annotation of the movie.
% individual results
Lastly, focussing on the audio-description, we compare results on the level of
individual subjects to results of a block-designed visual localizer experiment
done with the same individuals \citep{sengupta2016extension}.
% results
\todo[inline]{following paragraph will be similar to abstract, intro to
discussion and conclusion}
Our results demonstrate that (a), (b), (c).

@@

@@

\todo[inline]{subheadings are still in for better readability; structure could
be changed; at the moment its AO (PPA in primary contrast, robustness of PPA;
other clusters across contrasts), then the same for AV}

@@

\subsubsection{PPA in AO: primary contrasts}

@@

\subsubsection{PPA in AO: robustness across contrasts}

\todo[inline]{I consider tables of significant clusters in PPA for all contrasts
as overkill. Maybe in ``supporting material'' or reference to dataset? all sign.
clusters of all contrasts are in the comments}

% all eight AO PPA contrasts
In order to test the robustness of our findings, we created seven additional
$t$-contrasts for the AO stimulus (see Table \ref{tab:ao-contrasts}).
% results
Six of the overall eight AO contrasts show significant bilateral clusters in
anterior regions of the group PPA overlap (see Figure
% statement/answer
Results suggest that our findings do not depend on one specific contrasts but
are robust across differently designed
contrasts.\ref{fig:stability-slices}).\todo[inline]{Discussion: discuss that
contrasts make ``less and less'' sense}
% question
\todo[inline]{concluding statement often feels like interpretation/discussion}
% question 2

@@

\subsubsection{other sign. clusters across AO contrasts}
% precuneus = retrosplenial cortex ?
\todo[inline]{significant clusters in other regions than PPA could or should (?)
also be reported? Following is a short list; depends on the discussion I guess}
% ventral precuneus = retrosplenial cortex?
7 contrasts: bilateral clusters in RSC (and ventral precuneus);
% transverse occipital sulcus = occipital place area
6 contrasts (4 bilat.): transverse occipital sulcus (= OPA?);\todo{AO: clusters
are more lateral; AV: clusters are more medial}
% confounds
4 contrasts: right superior temporal gyrus; 3 contrasts: left superior temporal
gyrus; 3 contrasts: some (smallish) prefrontal clusters

\subsubsection{PPA in AV: primary contrasts}
% AV intro
Second as conceptual/cross-modal control of the group average results, we
analyzed data from the audio-visual movie by contrasting different kinds of
movie cuts independent(ly) from the frames' visual content following the cuts.
% AV primary contrast
The movie's primary $t$-contrast that compared cuts to an unfamiliar setting
with cuts within familiar setting (\texttt{vse\_new > vpe\_old}) yielded three
significant clusters (see Table \ref{tab:res-av-group1}, Figure
\ref{fig:group-slices}).
% PPA (others the clusters extend into).
One cluster spans across the midline and comprises (reported from posterior to
anterior), parts of the intracalcarine and cuneal cortex, the lingual gyrus and
retrosplenial cortex, the occipital and temporal fusiform gyrus and the
parahippocampal cortex in both hemispheres.
% LOC
Two additional homologous clusters are located in the superior lateral occipital
cortex (left bigger than right).

@@

\subsubsection{PPA in AV: robustness + other clusters across contrasts}

% all eight AV PPA contrasts
Here again in order to test the robustness of our findings, we created four
additional $t$-contrasts for the AV stimulus (see Table \ref{tab:av-contrasts}).
%% results: PPA
All of the overall five AV contrasts yielded significant bilateral clusters (in
that overlap with the PPA group mask/overlap [biggest overlap in posterior part
of group ROI = temporal occipital fusiform cortex] (see Figure
\ref{fig:stability-slices}).
% concluding statement
Here again, we can show that results can in principle be accomplished by varying
the design of the contrasts.
% question
\todo[inline]{concluding statement = interpretation/discussion?}

\subsection{negative controls in AO and AV}

\todo[inline]{contrasts are mentioned in the method section for the purpose of
    completeness. Shall results be reported (and later discussed)? Following is
a short overview}

% AO stimulus
AO stimulus (using annotation of movie cuts) all null results except: contrast 9
(vse\_new > pe\_old): right RSC.
% correlation of these nouns with cuts
Compare Figure \ref{fig:reg-corr}: nouns that cue change to a new setting are
correlated  with cuts to a new setting (vse\_new; r$\approx$0.3; second highest
correlation of a verbal cue category with a cut category.

% AV stimulus
AV stimulus. All contrasts using ``no cut condition'' > x yield null results.
The contrast that uses the narrator's nouns but in the AV stimulus (se\_new >
se\_old): right PPA, bilateral sup. lat. occ. cortex; LOC (l/r), right superior
parietal lobe. Compate Figure \ref{fig:reg-corr}: nouns that cue change to a new
setting are correlated  with cuts to a new setting (vse\_new; r$\approx$0.3;
second highest correlation of a verbal cue category with a cut category; AND:
nouns that cue change to a familiar setting are somewhat correlated with cuts to
a familiar setting (se\_old; r$\approx$0.4; highest correlation of a verbal
category with a cut category).

@@

\citep{baldassano2013differential} regarding DMN: the set of regions connected
more strongly to anterior PPA is strikingly similar to the Default Mode Network
(DMN) (Buckner et al., 2008; Fox et al., 2005; Raichle et al., 2001), which is
known to include the parahippocampal region.
%
The only portions of the DMN that do not show differential connectivity to
anterior PPA in our data are the rostral portion of the posterior cingulate
cortex (PCC) and the superior frontal cortex.  DMN has been implicated in a
large number of internally-focused tasks, one of its key roles involves
autobiographical memory (Buckner et al., 2008).  Models of recognition memory
have previously identified parahippocampal cortex as primarily encoding spatial
context information (Eichenbaum et al., 2007).
%
Data from Aminoff et al. (2007) has suggested that an anterior portion of PPA
may be involved in recall based on spatial context.
%
Our results are consistent with anterior PPA playing a more central role in
memory than posterior PPA, given anterior PPA's connectivity with the DMN. The
fact that anterior PPA had a lower sensitivity to our abstract object stimuli
does not necessarily im ply that this region does not use object information.
%
Previous work has shown that PPA responds to objects that have spatial
associations (Aminoff et al., 2007), are space-defining (Mullally and Maguire,
2011), and are navigationally-relevant (Janzen and Van Turennout, 2004).
%
These types of responses require spatial memory and cannot be based purely on
visual features like object shape.
%
If anterior PPA is involved in processing spatial context, then space-defining
or navigationally-relevant objects could activate anterior PPA more strongly
than our abstract objects, which were unfamiliar and provided no sense of
context or orientation.
%
Further experiments will be required to determine what type of object-related
information is used in this region.

@@

% contextual associations in anterior vs. posterior
\citep{aminoff2006parahippocampal}: We propose that the general function of the
PHC can be thought of as processing of contextual associations.
%
A subregion of the posterior PHC is particularly optimized to process spatially
organized contextual associations, whereas more anterior regions of the PHC are
optimized to process contextual associations in other, non-spatial domains.
% posterior PPA for visual visual associative information
Spatial contextual associations are a dominant property of places and scenes,
and thus scenes consistently activate this posterior subregion of the PHC (e.g.,
PPA; Epstein, 1998).
%
We do not argue that the PHC is not sensitive to spaces,
places, and spatial information, but rather that it is sensitive to spatial
stimuli because of the associative information (space-related associations) that
such stimuli evoke.

% Aminoff regarding contextual processing in the PHC and beyond
\citep{aminoff2006parahippocampal} regarding contextual associations processing
in PHC and beyond: We compared contextual processing of objects that are
strongly related to a specific place (e.g., an oven is found in a kitchen) with
contextual processing of objects that are highly associated with a context but
not a specific place (e.g., a birthday cake is found at a birthday party, but
the party can take place anywhere, at home, in a restaurant, at a park).
% method
Activity elicited by these two types of contextual objects was compared with the
activity elicited for objects that were only weakly associated with many
contexts (e.g., a lamp).
% results
Differential activity was found in the PHC and the RSC,
replicating the previous findings (Bar \& Aminoff, 2003, The
parahippocampal\dots).
% results 2
\citep{aminoff2006parahippocampal} Activity within the PHC, however, was
distributed along an anterior–posterior axis, with activity related to spatial
contexts focused in the posterior part of the PHC and activity evoked by
non-spatial contexts focused in the more anterior part of the PHC (Bar \&
Aminoff, 2003, Cortical analysis of visual context; Aminoff et al., 2007, The
parahippocampal cortex mediates spatial and nonspatial associations).
